---
title: "Traffic"
author: "Zhe Zhao(433707)"
date: "2022/5/24"
output:
  word_document: default
  html_document: default
---
## Introduction
The main target of this project is using regression methods of machine learning to build a model predicting the traffic on one of the highways for one-hourly intervals based on the training sample and generate predictions for all observations from the test sample.

```{r}
## import 
library(dplyr)
library(ggplot2)
library(caret)
library(lattice)
library(corrplot)
```


## Import data
Import data and check detailed information.
```{r}
## read data
traffic <- read.csv("C:/Users/zhaoz/Desktop/jiqixuexi/pro/traffic_train.csv")

## check detailed information
head(traffic)
str(traffic)
```

## Data Cleaning
"Traffic" is the dependent variable, so we need to remove the value equal to 0 to get the MAPE value, firstly we check if there are "0"s.
```{r}
## check if there is any "0"s in "traffic" 
any(traffic$traffic=="0")
```
Yes, there are "0" values, so We have to remove these rows with "traffic" equal to 0.
```{r}
## 0s transform into NA
traffic$traffic[which(traffic$traffic=="0")] <- NA

## omit NA value
traffic <- na.omit(traffic)

##check again
any(is.na(traffic))

```
Now there is no rows with "traffic" equal to 0.

Check outliers of variable temperature.
```{r}
## Find out if there are outliers
boxplot(traffic$temperature)
```

I also find there are some with temperature at absolute zero, we need to remove outliers.

We have to remove these rows with "temperature" equal to -273.1.
```{r}
## 0s transform into NA
traffic$temperature[which(traffic$temperature=="-273.1")] <- NA

## omit NA value
traffic <- na.omit(traffic)

##check again
any(is.na(traffic))
```
Now there are no outliers in variable temperature.

Lests set training sample as 70% and test sample as 30%.
```{r}
## set training sample and test sample 

set.seed(20030329) #set Random seed
traffic_which_training <- createDataPartition(traffic$traffic,
                          p = 0.7, list = FALSE)  #set 70% of data as training data
    
traffic_train <-traffic[c(traffic_which_training),] #training data
traffic_test <-traffic[-c(traffic_which_training),] #test data

## save data
save(list = c("traffic_train",
              "traffic_test"),
     file = "C:/Users/zhaoz/Desktop/jiqixuexi/pro/traffic_train_test.RData")

```
Check again if there are any NA values in training sample.
```{r}
## check if there are any NA values in training sample 
any(is.na(traffic_train))
```
No Na value in training sample.


Have a look at basic statistics of training sample and testing sample, We prefer that the results are similar.
```{r}

summary(traffic_train$traffic)
summary(traffic_test$traffic)
```
All these statistics are similar, training sample and testing sample could be acceptable.



## Data Exploration
In our experience, traffic is affected by time and we can visualize the result visually.
```{r}
plot(traffic_train$date_time,traffic_train$traffic )
```

We can see high traffic volume during the day and low traffic volume at night. 
p.s. date_time(hour)


Check outliers of variable rain_mm, .
```{r}
## Find out if there are outliers
boxplot(traffic_train$rain_mm)
```

Check outliers of variable snow_mm.
```{r}
# Find out if there are outliers
boxplot(traffic_train$snow_mm)
```

## Data Transform
Check the distribution of the dependent variable "traffic".
```{r}
# check the distribution (log transformation) of the dependent variable "traffic"
ggplot(traffic_train,
       aes(x = traffic)) +
  geom_histogram(fill = "blue",
                 bins = 100) +
  theme_bw()

```

Find character variables.
```{r}
## check character variables
traffic_train_character_vars <- 
  sapply(traffic_train, is.character) %>% 
  which() %>% 
  names()

## result
traffic_train_character_vars
```

There are two character variables, sort them.
```{r}
## sort 2 character variables
sapply(traffic_train[, traffic_train_character_vars], 
        function(x) 
          unique(x) %>% 
          length()) %>% 
  sort()

```
Transfer character variables to factor, find the levels of them.
```{r}
## transfer character variables to factor
traffic_train$weather_general <- factor(traffic_train$weather_general)
traffic_train$weather_detailed <- factor(traffic_train$weather_detailed)

## check levels of weather general and weather detailed
levels(traffic_train$weather_general)
levels(traffic_train$weather_detailed)
```

```{r}
## detailed information
table(traffic_train$weather_general)
plot(traffic_train$weather_general)
table(traffic_train$weather_detailed)
plot(traffic_train$weather_detailed)

## transfer int variables to num
traffic_train$date_time <- as.numeric(traffic_train$date_time)
traffic_train$clouds_coverage_pct <- as.numeric(traffic_train$clouds_coverage_pct)
```


Label variables weather_general and weather_detailed.
```{r}
## label weather_general
traffic_train$weather_general<-factor(traffic_train$weather_general,
                                           level=c("Clear","Clouds","Squall","Smoke","Fog","Thunderstorm","Haze",         
                                             "Drizzle","Snow", "Mist", "Rain"),
                                            labels = c(1:11))
```

```{r}
##  label weather_detailed
traffic_train$weather_detailed<-factor(traffic_train$weather_detailed,
                                                         level=c("broken clouds",  "drizzle",  "few clouds",
                                                                 "Mist",  "Rain","fog" ,"freezing rain", "haze" , "heavy                   intensity drizzle",   "heavy intensity rain", 
                "heavy snow" ,  "light intensity drizzle"            
              , "light intensity shower rain" ,        "light rain"                         
              , "light rain and snow",                 "light shower snow"                  
              ,"light snow"                                                      
              , "moderate rain",                       "overcast clouds"                    
              , "proximity shower rain" ,              "proximity thunderstorm"             
              , "proximity thunderstorm with drizzle", "proximity thunderstorm with rain"   
              , "scattered clouds",                    "shower drizzle"                     
              , "sky is clear" ,                       "smoke"                              
              , "snow" ,                             "squalls"                            
              , "thunderstorm"  ,                      "thunderstorm with drizzle"          
              , "thunderstorm with heavy rain"  ,      "thunderstorm with light drizzle"    
              , "thunderstorm with light rain" ,       "thunderstorm with rain"             
              , "very heavy rain"  ),
                                            labels = c(1:36))
```

```{r}
str(traffic_train)
```

Transfer variable weather_general and  weather_detailed into numeric variables, it will be easy to do correlation test.
```{r}
traffic_train$weather_detailed <- as.numeric(traffic_train$weather_detailed)
traffic_train$weather_general <- as.numeric(traffic_train$weather_detailed)

str(traffic_train)
```


Find numeric variables for correlation test.
```{r}
## find numeric variables
traffic_numeric_vars <- 
  # check if variable is numeric
  sapply(traffic_train, is.numeric) %>% 
  # select those which are
  which() %>% 
  # and keep just their names
  names()
## result
traffic_numeric_vars
```
Check correlations between variables, and draw correlation plot.
```{r}
# mutually correlated (irrelevant) variables
# calculate correlations between variables to identify redundant features

traffic_train_correlations <- 
  cor(traffic_train[, traffic_numeric_vars],
      use = "pairwise.complete.obs")
traffic_train_correlations

## result
corrplot(traffic_train_correlations)
```


```{r}
## correlation plot
corrplot.mixed(traffic_train_correlations,
               upper = "square",
               lower = "number",
               tl.col = "black", # color of labels (variable names)
               tl.pos = "lt")  # position of labels (lt = left and top)

```


Chose only variables "date_time", "traffic" and "temperature"

Creat new training sample with variables "date_time", "traffic" and "temperature" for next steps.
```{r}
traffic_train_pro <- traffic_train[,c("traffic", "date_time", "temperature")]

traffic_train_pro$date_time <- as.factor(traffic_train_pro$date_time)
```



## Linear regression model
```{r}
## Linear regression model 
traffic_lm1 <- lm(traffic ~ date_time + temperature, # formula
                 data = traffic_train_pro)
## output
summary(traffic_lm1)
```
Residuals
```{r}

head(traffic_lm1$residuals)

hist(traffic_lm1$residuals, breaks = 30)

# MSE is basically the average square residual

mean(traffic_lm1$residuals^2)

# MAE is the average absolute residual

mean(abs(traffic_lm1$residuals))

# similarly we can calculate Median Absolute Error

median(abs(traffic_lm1$residuals))
```

Write a simple function to summarize MSE, RMSE, MAE, MAPE, MedAE, MSLE and R2.
```{r}
# Write a simple function to summarize all the errors and R2

regressionMetrics <- function(real, predicted) {
  # Mean Square Error
  MSE <- mean((real - predicted)^2)
  # Root Mean Square Error
  RMSE <- sqrt(MSE)
  # Mean Absolute Error
  MAE <- mean(abs(real - predicted))
  # Mean Absolute Percentage Error
  MAPE <- mean(abs(real - predicted)/real)
  # Median Absolute Error
  MedAE <- median(abs(real - predicted))
  # Mean Logarithmic Absolute Error
  MSLE <- mean((log(1 + real) - log(1 + predicted))^2)
  # R2
  R2 <- cor(predicted, real)^2

  result <- data.frame(MSE, RMSE, MAE, MAPE, MedAE, MSLE, R2)
  return(result)
}

# lets apply it to our model
regressionMetrics(real = traffic_train_pro$traffic,
                  predicted = predict(traffic_lm1))
```

R^2 = 0.7701107 and MAPE = 1.164449, seems okay, so lets check how it's performance in testing sample.
Testing sample is names "traffic_test", separated from "traffic_train".
```{r}
## Linear regression model for testing data
traffic_test$date_time <- as.factor(traffic_test$date_time)
traffic_test$predicted <- traffic_lm1 %>% predict(traffic_test)

## output
print(bquote(R^2==.(R2(traffic_test$predicted,traffic_test$traffic))))
```
We can check the MAPE value, it is even lower.
We can also have a look at the value of R^2 = 0.767948620226271, still high!


## Conslusion
I built SVR and KNN model also, but the results are similar or worse, so i just chose Linear Regression model. The value of R square and MAPE Linear Regression model are not very good, but they are still better than the other models I used. 
