---
title: "Traffic"
author: "Zhe Zhao(433707)"
date: "2022/5/24"
output: html_document
---

## Introduction
The main target of this project is using regression methods of machine learning to build a model predicting the traffic on one of the highways for one-hourly intervals based on the training sample and generate predictions for all observations from the test sample.

```{r}
library(dplyr)
library(ggplot2)
library(caret)
library(lattice)
library(corrplot)
```

## Import data
Import data and check detailed information.
```{r}
## read data
traffic <- read.csv("C:/Users/zhaoz/Desktop/jiqixuexi/pro/traffic_train.csv")
```

"Traffic" is the dependent variable, so we need to remove the value equal to 0 to get the MAPE value, firstly we check if there are "0"s.
There are "0" values, so We have to remove these rows with "traffic" equal to 0.
```{r}
## 0s transform into NA
traffic$traffic[which(traffic$traffic=="0")] <- NA
## omit NA value
traffic <- na.omit(traffic)
```

```{r}
## 0s transform into NA
traffic$temperature[which(traffic$temperature=="-273.1")] <- NA

## omit NA value
traffic <- na.omit(traffic)
```

Set training sample as 70% and testing sample as 30%.
```{r}
## set training sample and test sample 

set.seed(20030329) #set Random seed
traffic_which_training <- createDataPartition(traffic$traffic,
                          p = 0.7, list = FALSE)  #set 70% of data as training data
    
traffic_train <-traffic[c(traffic_which_training),] #training data
traffic_test <-traffic[-c(traffic_which_training),] #test data

## save data
save(list = c("traffic_train",
              "traffic_test"),
     file = "C:/Users/zhaoz/Desktop/jiqixuexi/pro/traffic_train_test.RData")

```

I chose only 3 variables: "traffic", "date_time" and "temperature".
```{r}
traffic_train_pro <- traffic_train[,c("traffic", "date_time", "temperature")]

traffic_train_pro$date_time <- as.factor(traffic_train_pro$date_time)
```

## Linear regression model
Why linear regression?
1. This modeling speed is fast, and it does not require very complicated calculations. It still runs very fast in the case of a large amount of data, especially there are almost 30,000 pieces of data.
2. It could give us an explanation for each variable in terms of coefficients.

Why not KNN(K-Nearest Neighbor) or SVR(support vector regression)?
The computational complexity are high, the amount of computation are too large, so they took really long time.
However, the MAPE value of LR and SVR(linear kernel) are really similar,  LNN has Worse MAPE. 

All in all, speed of LR is fast and lower MAPE.
```{r}
## Linear regression model 
traffic_lm1 <- lm(traffic ~ date_time + temperature, # formula
                 data = traffic_train_pro)
## output
summary(traffic_lm1)
```
Residuals
```{r}

head(traffic_lm1$residuals)

hist(traffic_lm1$residuals, breaks = 30)

# MSE is basically the average square residual

mean(traffic_lm1$residuals^2)

# MAE is the average absolute residual

mean(abs(traffic_lm1$residuals))

# similarly we can calculate Median Absolute Error

median(abs(traffic_lm1$residuals))
```
Write a simple function to summarize MSE, RMSE, MAE, MAPE, MedAE, MSLE and R2.
```{r}
# Write a simple function to summarize all the errors and R2

regressionMetrics <- function(real, predicted) {
  # Mean Square Error
  MSE <- mean((real - predicted)^2)
  # Root Mean Square Error
  RMSE <- sqrt(MSE)
  # Mean Absolute Error
  MAE <- mean(abs(real - predicted))
  
  # Mean Absolute Percentage Error!!
  MAPE <- mean(abs(real - predicted)/real)
  
  # Median Absolute Error
  MedAE <- median(abs(real - predicted))
  # Mean Logarithmic Absolute Error
  MSLE <- mean((log(1 + real) - log(1 + predicted))^2)
  # R2
  R2 <- cor(predicted, real)^2

  result <- data.frame(MSE, RMSE, MAE, MAPE, MedAE, MSLE, R2)
  return(result)
}

# apply it to LR model
regressionMetrics(real = traffic_train_pro$traffic,
                  predicted = predict(traffic_lm1))
```

R^2 = 0.7701107 and MAPE = 1.164449, seems okay, so lets check how it's performance in testing sample.
Testing sample is names "traffic_test", separated from "traffic_train".
```{r}
## Linear regression model for testing data
traffic_test$date_time <- as.factor(traffic_test$date_time)
traffic_test$predicted <- traffic_lm1 %>% predict(traffic_test)

## output
print(mean(abs(traffic_test$traffic - traffic_test$predicted)/traffic_test$traffic))
print(bquote(R^2==.(R2(traffic_test$predicted,traffic_test$traffic))))
```
We can check the MAPE value, it is even lower.
We can also see the value of R^2 = 0.767948620226271, still high!


## Conslusion
I built SVR and KNN model also, but the results are similar or worse, so i just chose Linear Regression model. The value of R square and MAPE Linear Regression model are not very good, but they are still better than the other models I used. 
